\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\hypersetup{colorlinks=true,urlcolor=blue,linkcolor=black}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\begin{document}

\begin{center}
{\Large\bfseries Boundary Information Geometry for S-box Side-Channel Leakage:
Walsh Covariance, Active Fisher Structure, and CPA Belief Dynamics}
\end{center}

\vspace{1em}

\textbf{Authors and affiliations}

\begin{enumerate}
  \item \textbf{St\'ephane Ga\"el R. Ekodeck} (corresponding author)\\
        ORCID: \href{https://orcid.org/0000-0002-8094-8832}{0000-0002-8094-8832}\\
        University of Yaounde I, LIRIMA, Team GRIMCAPE, Yaounde, Cameroon\\
        IRD, UMI, UMMISCO, IRD France Nord, Bondy 93143, France\\
        Sorbonne Universites, Univ. Paris 06, UMI 209, UMMISCO, Paris 75005, France\\
        Email: \href{mailto:ekodeckstephane@gmail.com}{ekodeckstephane@gmail.com}

  \item \textbf{Serge Alain Ebele}\\
        ORCID: \href{https://orcid.org/0000-0001-7036-7068}{0000-0001-7036-7068}\\
        University of Yaounde I, LIRIMA, Team GRIMCAPE, Yaounde, Cameroon\\
        IRD, UMI, UMMISCO, IRD France Nord, Bondy 93143, France\\
        Sorbonne Universites, Univ. Paris 06, UMI 209, UMMISCO, Paris 75005, France\\
        Email: \href{mailto:serge-alain.ebele@facsciences-uy1.cm}{serge-alain.ebele@facsciences-uy1.cm}

  \item \textbf{Prosper Rosaire Mama Assandje}\\
        ORCID: \href{https://orcid.org/0000-0002-1234-5678}{0000-0002-1234-5678}\\
        Department of Mathematics, University of Yaounde I, Yaounde, Cameroon\\
        Email: \href{mailto:mamarosaire@facsciences-uy1.cm}{mamarosaire@facsciences-uy1.cm}

  \item \textbf{Rene Ndoundam}\\
        ORCID: \href{https://orcid.org/0000-0003-1105-762X}{0000-0003-1105-762X}\\
        University of Yaounde I, LIRIMA, Team GRIMCAPE, Yaounde, Cameroon\\
        IRD, UMI, UMMISCO, IRD France Nord, Bondy 93143, France\\
        Sorbonne Universites, Univ. Paris 06, UMI 209, UMMISCO, Paris 75005, France\\
        Email: \href{mailto:rndoundam@gmail.com}{rndoundam@gmail.com}
\end{enumerate}

\vspace{1em}
\textbf{Corresponding author details}

Dr. St\'ephane Ga\"el R. Ekodeck\\
University of Yaounde I, LIRIMA, Team GRIMCAPE, Yaounde, Cameroon\\
Email: \href{mailto:ekodeckstephane@gmail.com}{ekodeckstephane@gmail.com}

\vspace{1em}
\textbf{Declarations}

\textbf{Competing interests}

The authors declare that they have no competing interests.

\textbf{Authors contributions}

Conceptualization and methodology: S.G.R.E. and S.A.E.; formal analysis:
S.G.R.E., S.A.E., P.R.M.A., and R.N.; software and reproducible computations:
S.G.R.E.; validation: S.A.E., P.R.M.A., and R.N.; supervision and project
administration: R.N.; writing--original draft: S.G.R.E.; writing--review and
editing: S.G.R.E., S.A.E., P.R.M.A., and R.N. All authors read and approved
the final manuscript.

\textbf{Acknowledgements}

This work was supported by UMMISCO, LIRIMA, and the University of Yaounde I,
Cameroon. The authors used AI-assisted language tools to improve readability.
After using these tools, the authors reviewed and edited the content and take
full responsibility for the publication.

\textbf{Authors' information}

\textbf{St\'ephane Ga\"el R. Ekodeck} is a senior lecturer and researcher in
computer science at the University of Yaounde I, Cameroon, affiliated with
LIRIMA, Team GRIMCAPE, and UMMISCO. His research interests include information
security, steganography, privacy-preserving data processing, statistical
modelling, and secure information systems. His recent work focuses on
mathematical and information-theoretic models for secure multimedia
communication, coverless steganography, and geometry-aware optimization for
generative systems.

\textbf{Serge Alain Ebele} is a senior lecturer and researcher in computer
science at the University of Yaounde I, Cameroon, affiliated with LIRIMA, Team
GRIMCAPE, and UMMISCO. His research interests include complex systems,
information security, discrete dynamical systems, cryptographic and
steganographic models, and mathematical approaches to secure computing.

\textbf{Prosper Rosaire Mama Assandje} is an assistant lecturer and researcher
in mathematics and information geometry affiliated with the University of
Yaounde I, Cameroon. His research interests include information geometry,
Hessian geometry, statistical manifolds, integrable gradient systems,
Hamiltonian systems, and geometric methods for learning.

\textbf{Rene Ndoundam} is a full professor and researcher in computer science
at the University of Yaounde I, Cameroon, affiliated with LIRIMA, Team
GRIMCAPE, and UMMISCO. His research interests include theoretical computer
science, cryptography, steganography, discrete dynamical systems, neural
recurrence equations, distributed systems, and information security.

\vspace{1em}
\textbf{Article type}

Research Article

\end{document}
