% ============================================================================
% SECTION 7: NUMERICAL COMPUTATIONS
% ============================================================================
¥section{Reproducible Security Diagnostics}¥label{sec:numerical}

The computations in this section connect the closed-form geometry to
attack-relevant side-channel diagnostics.  First, they verify the exact
boundary covariance and active-eigenspace formulas on standard lightweight
S-boxes.  Second, they separate quantities that are structural invariants from
quantities that influence CPA discrimination between key hypotheses.  Third,
they validate the pairwise-gap interpretation through a reproducible Monte
Carlo CPA experiment.

¥subsection{Target S-boxes}¥label{subsec:sboxes}

We evaluate the 4-bit bijective S-boxes used in PRESENT, GIFT, and SKINNY
¥cite{bogdanov2007present,banik2017gift,beierle2016skinny}:
¥begin{align*}
S_{¥mathrm{PRESENT}}&=[C,5,6,B,9,0,A,D,3,E,F,8,4,7,1,2],¥¥
S_{¥mathrm{GIFT}}&=[1,A,4,C,6,F,3,9,2,D,B,7,5,0,8,E],¥¥
S_{¥mathrm{SKINNY}}&=[C,6,9,0,1,A,2,B,3,8,5,D,4,E,7,F].
¥end{align*}
These examples are useful because they are widely used in lightweight
cryptography and because exhaustive enumeration over all inputs, masks, and
key hypotheses is possible.

¥subsection{Walsh spectra and boundary covariance}¥label{subsec:wht-spectra}

For each S-box we compute the full WHT table
¥[
  ¥widehat S(¥alpha,¥beta)=¥sum_{a¥in¥B{4}}
  (-1)^{¥alpha¥cdot a¥oplus¥beta¥cdot S(a)}.
¥]
The three tested S-boxes have the same absolute Walsh-coefficient distribution
over nonzero masks:
¥[
  ¥#|¥widehat S|=0:123,¥qquad
  ¥#|¥widehat S|=4:96,¥qquad
  ¥#|¥widehat S|=8:36.
¥]
Therefore diagnostics depending only on the Walsh amplitudes coincide for
these three examples.  This shared spectral profile provides a useful baseline
for reading the later CPA-specific quantities.

¥begin{figure}[t]
¥centering
¥includegraphics[width=¥textwidth]{fig1_wht_spectra.pdf}
¥caption{Walsh spectra. The plotted Walsh tables show the same amplitude
profile for PRESENT, GIFT, and SKINNY, which explains why purely
amplitude-based diagnostics coincide for these examples.}
¥label{fig:wht-spectra}
¥end{figure}

¥begin{figure}[t]
¥centering
¥includegraphics[width=0.74¥textwidth]{fig2_wht_histogram.pdf}
¥caption{Walsh amplitude distribution. The histogram records the shared counts
of absolute Walsh coefficients over nonzero masks for the three tested S-boxes.}
¥label{fig:wht-histogram}
¥end{figure}

The boundary covariance matrix is built from Proposition‾¥ref{thm:fim}.  For
$s=4$ it is a $255¥times255$ positive semidefinite matrix.  Consistent with
Theorem‾¥ref{thm:active-rank}, all three matrices have rank $15$ and active
eigenvalues equal to $16$ up to numerical precision.

¥begin{table}[t]
¥centering
¥caption{Boundary covariance structure for the tested 4-bit bijections.}
¥label{tab:cov-rank}
¥begin{tabular}{lccc}
¥toprule
S-box & Matrix size & Rank & Active eigenvalues ¥¥
¥midrule
PRESENT & $255¥times255$ & $15$ & $16$ repeated $15$ times ¥¥
GIFT    & $255¥times255$ & $15$ & $16$ repeated $15$ times ¥¥
SKINNY  & $255¥times255$ & $15$ & $16$ repeated $15$ times ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

¥subsection{Local spectral separation}¥label{subsec:fr-distances}

Using‾¥eqref{eq:fr-local}, we compute local diagonal Walsh/Fisher separation
proxies between key hypotheses.  This local covariance-weighted coordinate
displacement is evaluated in the regularized Walsh/Fisher coordinates
introduced earlier.  The key-translation rule implies that the value depends
only on the key difference $k¥oplus k'$.

¥begin{figure}[t]
¥centering
¥includegraphics[width=0.8¥textwidth]{fig3_fr_distance_matrix.pdf}
¥caption{Local key-separation proxy. The matrix visualises the
covariance-weighted Walsh displacement between key hypotheses in the local
Walsh/Fisher coordinates.}
¥label{fig:local-distance}
¥end{figure}

¥begin{figure}[t]
¥centering
¥includegraphics[width=0.74¥textwidth]{fig5_fr_vs_hw.pdf}
¥caption{Local proxy versus key-difference Hamming weight.  The plot aggregates
the same covariance-weighted Walsh displacement by the Hamming weight of
$k¥oplus k'$.}
¥label{fig:local-distance-hw}
¥end{figure}

¥begin{table}[t]
¥centering
¥caption{Interpretation of the key-separation quantities.}
¥label{tab:distance-status}
¥begin{tabular}{ll}
¥toprule
Quantity & Status ¥¥
¥midrule
$¥KL(P_k¥|P_{k'})$ for $k¥neq k'$ & Infinite, because supports are disjoint ¥¥
$d_F(¥delta_k,¥delta_{k'})$ on the belief simplex & Exactly $¥pi$ ¥¥
$d_{¥mathrm{loc}}^2(k,k')$ in‾¥eqref{eq:fr-local} & Local Walsh/Fisher covariance proxy ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

¥subsection{CPA leakage gaps}¥label{subsec:pairwise-gap-numerical}

For bijective 4-bit S-boxes and uniform inputs, the scalar Hamming-weight
variance in Proposition‾¥ref{thm:per-trace} is always
¥[
  I_S(¥HW,¥sigma)=¥frac{1}{¥sigma^2}.
¥]
This scalar is a baseline leakage descriptor shared by PRESENT, GIFT, and
SKINNY.  Pairwise gaps
¥[
  ¥Delta(¥delta)=¥frac1{16}¥sum_{a¥in¥B{4}}
  ¥bigl(¥HW(S(a))-¥HW(S(a¥oplus¥delta))¥bigr)^2,
  ¥qquad ¥delta¥neq0,
¥]
are more informative for CPA because the expected log-Bayes gain against a
wrong key with difference $¥delta$ is $¥Delta(¥delta)/(2¥sigma^2)$.

¥begin{table}[t]
¥centering
¥caption{Pairwise Hamming-weight leakage gaps over nonzero key differences.}
¥label{tab:pairwise-gaps}
¥begin{tabular}{lccc}
¥toprule
S-box & $¥min_{¥delta¥neq0}¥Delta(¥delta)$ &
$¥max_{¥delta¥neq0}¥Delta(¥delta)$ &
$¥frac1{15}¥sum_{¥delta¥neq0}¥Delta(¥delta)$ ¥¥
¥midrule
PRESENT & $1$ & $7/2$ & $32/15$ ¥¥
GIFT    & $5/4$ & $7/2$ & $32/15$ ¥¥
SKINNY  & $1$ & $3$   & $32/15$ ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

Table‾¥ref{tab:pairwise-gaps} shows how the scalar variance descriptor and the
pairwise structure play complementary roles: all three S-boxes have the same
average pairwise gap, while the worst and best key-difference cases vary.
These pairwise gaps are the
quantities that enter expected posterior-odds growth in the Gaussian leakage
model.

¥begin{table}[t]
¥centering
¥caption{Operational CPA interpretation of the pairwise gaps.  The listed
deltas are nonzero key differences $¥delta=k¥oplus k^*$ in hexadecimal.}
¥label{tab:gap-distribution}
¥scriptsize
¥setlength{¥tabcolsep}{2pt}
¥begin{tabular}{@{}p{0.13¥textwidth}p{0.17¥textwidth}p{0.17¥textwidth}p{0.43¥textwidth}@{}}
¥toprule
S-box & Hardest gaps $¥min¥Delta$ & Easiest gaps $¥max¥Delta$ &
Gap multiset over $¥delta¥neq0$ ¥¥
¥midrule
PRESENT & $1$ at $¥{A,E¥}$ & $7/2$ at $¥{6¥}$ &
$1^2$; $(5/4)^1$; $(3/2)^2$; $(7/4)^3$; $(5/2)^2$;
$(11/4)^1$; $3^2$; $(13/4)^1$; $(7/2)^1$ ¥¥
GIFT & $5/4$ at $¥{C,E,F¥}$ & $7/2$ at $¥{7¥}$ &
$(5/4)^3$; $(3/2)^2$; $2^2$; $(9/4)^3$; $(5/2)^2$;
$3^2$; $(7/2)^1$ ¥¥
SKINNY & $1$ at $¥{5¥}$ & $3$ at $¥{B¥}$ &
$1^1$; $(5/4)^1$; $(3/2)^1$; $(7/4)^1$; $2^3$;
$(9/4)^2$; $(5/2)^3$; $(11/4)^2$; $3^1$ ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

Table‾¥ref{tab:gap-distribution} gives the CPA-oriented reading of the
pairwise theory.  PRESENT and SKINNY have a smaller hardest-key gap than GIFT
under this leakage model, so the model-based estimate
¥eqref{eq:gap-sample-estimate} assigns a larger trace budget to the hardest
posterior-odds margin for those S-boxes.  The comparison is local to the
specified S-box computation and leakage model, which is the level at which CPA
key hypotheses are evaluated.

¥subsection{Monte Carlo CPA validation}¥label{subsec:cpa-simulation}

To connect the gap analysis to an attack observable, we simulate the Bayesian
CPA model of Section‾¥ref{sec:cpa}.  For each S-box, we use true key
$k^*=0$, uniformly sampled plaintext nibbles, leakage
$W=¥HW(S(a¥oplus k^*))+¥varepsilon$, Gaussian noise standard deviation
$¥sigma=2$, and $3000$ Monte Carlo trials with seed $20260622$.  For every
trace prefix, all 16 key log-likelihoods are accumulated and the true-key rank
is recorded.  The script ¥texttt{simulate¥_cpa.py} regenerates both
Fig.‾¥ref{fig:cpa-simulation} and the CSV summary
¥texttt{cpa¥_simulation¥_results.csv}.

¥begin{figure}[t]
¥centering
¥includegraphics[width=¥textwidth]{fig6_cpa_simulation.pdf}
¥caption{Simulated CPA under Hamming-weight Gaussian leakage.  The left panel
shows empirical success probability and the right panel shows mean true-key
rank over $3000$ trials.  The curves translate the pairwise-gap analysis into
trace-dependent attack observables.}
¥label{fig:cpa-simulation}
¥end{figure}

The simulation confirms the qualitative role of the pairwise gaps.  At
$64$ traces, the empirical success probabilities are approximately $0.923$ for
PRESENT, $0.950$ for GIFT, and $0.947$ for SKINNY; at $128$ traces, all three
success probabilities are between $0.993$ and $0.996$.  Mean true-key ranks
decrease toward one for all three S-boxes.  The finite-trace curves summarize
the combined effect of all wrong-key competitors, while the minimum gap in
Table‾¥ref{tab:pairwise-gaps} controls the hardest pairwise posterior-odds
growth.

¥begin{figure}[t]
¥centering
¥includegraphics[width=0.74¥textwidth]{fig4_per_trace_info.pdf}
¥caption{Per-trace leakage baseline. The scalar Hamming-weight variance is the
same for all bijective 4-bit S-boxes under uniform inputs, so CPA discrimination
is completed by pairwise leakage gaps that determine key-hypothesis
discrimination.}
¥label{fig:per-trace-info}
¥end{figure}

¥subsection{Active-eigenspace curvature}¥label{subsec:curv-numerical}

The curvature diagnostic of Section‾¥ref{sec:curvature} is evaluated in the
15-dimensional active eigenspace of the boundary covariance matrix.  The result
is constant for all coordinate planes in the tested cases.

¥begin{table}[t]
¥centering
¥caption{Projected active-eigenspace curvature diagnostic for the tested 4-bit S-boxes.}
¥label{tab:curvatures}
¥begin{tabular}{lcccc}
¥toprule
S-box & Active dimension & Planes & Min $K$ & Max $K$ ¥¥
¥midrule
PRESENT & $15$ & $105$ & $-0.25$ & $-0.25$ ¥¥
GIFT    & $15$ & $105$ & $-0.25$ & $-0.25$ ¥¥
SKINNY  & $15$ & $105$ & $-0.25$ & $-0.25$ ¥¥
¥bottomrule
¥end{tabular}
¥end{table}

¥subsection{Reproducibility checklist}¥label{subsec:repro-checklist}

The computational claims in this section are reproducible from the additional
files.  The scripts ¥texttt{gen¥_figures.py}, ¥texttt{simulate¥_cpa.py},
¥texttt{compute¥_curvatures.py}, and ¥texttt{verify¥_curvature.py} regenerate
the figures, CPA simulation results, active covariance diagnostics, and
curvature checks.  The
rank/eigenvalue statement is proved independently in
Theorem‾¥ref{thm:active-rank}; the scripts provide finite checks for the
reported 4-bit examples.
